---
title: "Homework 6"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(modelr)
library(mgcv)
library(p8105.datasets)
```

## Problem 1

Adapted solution for Problem 1 included below.

```{r weather_tidy, message = FALSE, cache = TRUE}
weather_df = rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(name = recode(id, USW00094728 = "CentralPark_NY"),
         tmin = tmin / 10,
         tmax = tmax / 10)
```

```{r weather_model_plot}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(models = map(strap, ~lm(tmax ~ tmin, data = .x)),
         results = map(models, broom::glance)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  ggplot(aes(x = r.squared)) + geom_density()
```

```{r weather_bootstrap_plot}
weather_df %>% 
  modelr::bootstrap(n = 1000) %>% 
  mutate(models = map(strap, ~lm(tmax ~ tmin, data = .x)),
         results = map(models, broom::tidy)) %>% 
  select(-strap, -models) %>% 
  unnest(results) %>% 
  select(id = `.id`, term, estimate) %>% 
  pivot_wider(names_from = term, 
              values_from = estimate) %>% 
  mutate(log_b0b1 = log(`(Intercept)` * tmin)) %>% 
  ggplot(aes(x = log_b0b1)) + geom_density()
```

## Problem 2

We first import and tidy `homicides` data from The Washington Post.

```{r hm_tidy, warning = FALSE, message = FALSE, show_col_types = FALSE}
homicides = read_csv("data/homicide-data.csv") %>%
  mutate(city_state = str_c(city, ", ", state),
         sol_hm = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(victim_age),
         victim_race = fct_relevel(as.factor(victim_race), "White"),
         victim_sex = fct_relevel(as.factor(victim_sex), "Male")) %>%
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")), 
         victim_race %in% c("Black", "White"),
         victim_sex != "Unknown") %>%
  relocate(city_state, .after = state)
```

We proceed to fit a logistic regression model, with case resolution status (i.e. `sol_hm`, a binary variable) as the outcome, and `victim_age`, `victim_race`, and `victim_sex` as the independent variables. We then report the adjusted odds ratio and corresponding confidence interval for the effect of being a female victim, versus male victims, on likelihood of case resolution, holding other predictors constant.

```{r hm_glm_balt}
balt_fit_logistic = homicides %>%
  filter(city_state == "Baltimore, MD") %>%
  glm(sol_hm ~ victim_age + victim_race + victim_sex, data = ., family = binomial())

balt_fit_logistic %>% 
  broom::tidy(conf.int = TRUE) %>% 
  mutate(or = exp(estimate),
         conf_low = exp(conf.low),
         conf_high = exp(conf.high)) %>%
  filter(term == "victim_sexFemale") %>%
  select(c(1, 8:10)) %>% 
  knitr::kable(digits = 3)
```

We fit logistic regression models for each city and then again report the estimated odds ratio and corresponding confidence intervals for the effect of being a female victim on the probability of case resolution, accounting for other predictors specified.

```{r hm_glm_all_cities}
cities_log_reg = homicides %>%
  nest(data = -city_state) %>%
  mutate(models = map(data, ~glm(sol_hm ~ victim_age + victim_race + victim_sex, data = ., family = binomial())),
         results = map(models, broom::tidy, conf.int = TRUE)) %>%
  select(-data, -models) %>%
  unnest(results) %>%
  mutate(or = exp(estimate),
         conf_low = exp(conf.low),
         conf_high = exp(conf.high)) %>%
  filter(term == "victim_sexFemale") %>%
  select(c(1:2, 9:11))
```

Finally, we plot estimated odds ratios and confidence intervals by city, in ascending order of ORs. Among 47 cities, New York, NY has the highest odds ratio for the effect of being a female victim on likelihood of the homicide being solved, while New York; Long Beach, CA; and San Bernardino, CA, have the largest confidence intervals for their corresponding ORs. 

```{r hm_plot_or_ci, fig.width = 9, fig.asp = 0.6}
cities_log_reg %>% 
  mutate(city_state = fct_reorder(city_state, or)) %>%
  ggplot(aes(x = city_state, y = or, color = city_state)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high)) +
  labs(title = "Estimated adjusted odds ratios and confidence intervals for solved homicides, female versus male victims, by city",
       x = "City",
       y = "Estimated odds ratios (with confidence intervals)") +
  theme(legend.position = "none",
        plot.title = element_text(size = 11),
        axis.text.x = element_text(angle = 45, hjust = 1))
```

## Problem 3

We import, tidy, and check for missing values in `birthweights` data.

```{r bw_tidy, message = FALSE, show_col_types = FALSE}
birthweights = read_csv("data/birthweight.csv") %>%
  mutate(babysex = as.factor(babysex),
         frace = as.factor(frace),
         mrace = as.factor(mrace),
         malform = as.factor(malform),
         fincome = fincome * 100)

birthweights %>%
  skimr::skim()
```

We propose a regression model (i.e. `mod_a`) with the following predictors that are hypothesized to be associated with birthweight (i.e. `bwt`): weight of the mother at delivery (i.e. `delwt`), family income (i.e. `fincome`), gestational age (i.e. `gaweeks`), existence of malformations (i.e. `malform`), age of the mother at delivery (i.e. `momage`), and average daily count of cigarettes consumed during pregnancy (i.e. `smoken`).

```{r bw_reg}
mod_a = lm(bwt ~ delwt + fincome + gaweeks + malform + momage + smoken, data = birthweights)

mod_a %>%
  broom::tidy() %>%
  knitr::kable(digits = 3)
```

We proceed in our modeling process by plotting residuals versus predicted values for `mod_a`. 

```{r bw_plot_resid}
birthweights %>% 
  add_residuals(mod_a, "resid") %>% 
  add_predictions(mod_a, "fitted") %>% 
  ggplot(aes(x = fitted, y = resid)) + 
  geom_point() +
  labs(title = "Residuals versus predicted values for mod_a",
       x = "Predicted values",
       y = "Residuals")
```

Next, we use cross-validation to compare the predictive performance of `mod_a` with two alternative models: `mod_b`, which regresses birthweight on length of the baby at birth (i.e. `blength`) and gestational age; and `mod_c`, which regresses birthweight on head circumference of the baby at birth (i.e. `bhead`), length of the baby at birth, sex of the baby (i.e. `babysex`), and interactions.

```{r bw_comp_cv}
cv_df = crossv_mc(birthweights, 100) %>% 
  mutate(train = map(train, as_tibble),
         test = map(test, as_tibble))

cv_df = cv_df %>% 
  mutate(mod_a = map(train, ~lm(bwt ~ delwt + fincome + gaweeks + malform + momage + smoken, data = .x)),
         mod_b = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
         mod_c = map(train, ~lm(bwt ~ bhead + blength + babysex + 
                                      bhead * blength + 
                                      blength * babysex +
                                      bhead * babysex +
                                      bhead * blength * babysex, 
                               data = .x)),
    rmse_mod_a = map2_dbl(mod_a, test, ~rmse(model = .x, data = .y)),
    rmse_mod_b = map2_dbl(mod_b, test, ~rmse(model = .x, data = .y)),
    rmse_mod_c = map2_dbl(mod_c, test, ~rmse(model = .x, data = .y)))
```

```{r bw_comp_plot, fig.width = 9, fig.asp = 0.6}
cv_df %>% 
  select(c(7:9)) %>% 
  pivot_longer(everything(), names_to = "model", values_to = "rmse", names_prefix = "rmse_") %>% 
  mutate(model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse)) + geom_violin() +
  labs(title = "Predictive performance of regression models for birthweight",
       x = "Model",
       y = "Root mean squared error")
```
